### 2022-09-27 chapters/optimization/python-dir/python.tex
update : multiprocessing p1

As one can notice in~\Cref{fig:pythontreatement}, the built-in methods are the most energy-saving ones, while the customize \texttt{while} loop is the heaviest.
Moreover, when increasing the size of the collection, the map with lambda functions tends to consume less energy than the predetermined one.
The reason behind such behavior is that Python treats these functions as local variables, unlike the predefined ones, which are global in our case. Therefore, they are faster and consume less energy.

\paragraph*{subsubsection}{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python application is to follow the guidelines and to privilege the built-in functions.

\subsubsection{Python \& Multiprocessing}

% \subsubsection{Cost of Parallelism in Python}
Additionally, \Cref{fig:python_multiprocessing} reports on the energy consumption of Python applications when introducing parallelism.
Python applications are single-threaded systems constrained by the \emph{global lock system} (GLS).
However, due to the increase of cores/threads per CPU, many Python libraries started to take advantage of this hardware feature by allowing concurrent execution of Python processes.
For example, the multiprocessing library\footnote{\url{https://docs.python.org/3/library/multiprocessing.html}} spawns sub-processes to increase the degree of concurrency of Python applications.
As one can see in \Cref{fig:python_multiprocessing}, the energy consumption is correlated with the number of threads until reaching the limit of physical cores, when concurrent processes start to compete for the CPU.

@@

\input{python_interperters.tex}
% \subsection{threads to validity \note{missing}}
\clearpage
\section{Summary}
Because Python is widely used by software services deployed in public and private cloud infrastructures, increasing the energy efficiency of Python-based applications will reduce ICT carbon emissions.
This chapter discusses various ways for increasing the energy efficiency of Python-based applications. It first explains why such a choice was made, and then it develops a method to enhance the energy efficiency of the most commonly employed use cases, which are revealed to include web development and machine learning.
In web development, we found that the majority of the energy is wasted while processing requests, therefore adopting the proper strategy can reduce energy consumption by tenfold.
Following that, we looked at the energy usage for the machine learning usecase and discovered that it is not proportional to model accuracy.
As a result, a reduction in accuracy can result in large energy savings.
Following that, we dug deeper to investigate how data structures, parallelism, and iterative methods affect energy use.
In the second section, we presented a simple technique for optimizing the energy usage of Python-based apps. without making substantial changes to the code This is the use of many Python interpreters
Section 2 started by categorizing and filtering these implementations into three major kinds (compiler, interpreter, and extra libraries).
Then we conducted a series of experiments to compare the energy consumption of various interpreters.
The findings indicate the lack of a general solution and the importance of selecting the appropriate interpreter for the appropriate use case. In certain circumstances, some of these answers may be the best, while in others, they may be the worst.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.
We believe that our contributions will benefit a wide spectrum of legacy systems, not only reducing ICT's carbon footprint, but also lowering their cloud bills for the resources these services need.

>>>>

As one can notice in~\Cref{fig:pythontreatement}, the built-in methods are the most energy-saving ones, while the customize \texttt{while} loop is the heaviest.
Moreover, when increasing the size of the collection, the map with lambda functions tends to consume less energy than the predefined one.
The reason behind such behavior is that Python treats these functions as local variables, unlike the predefined ones, which are global in our case. Therefore, they are faster and consume less energy.

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python application is to follow the guidelines and to privilege the built-in functions.

\subsubsection{Python \& Multiprocessing}
The purpose of this part is to study the impact of concurrency on the energy consumption for Python applications.
To do so, we run a simple code snippet that computes the sum of the first $N$ numbers using the standard concurrency libraries \texttt{multiprocessing} and \texttt{multihtreading}. Using Desktop  machine with Intel(R) Core(TM) i7-4800MQ CPU @ 2.70GHz with four cores and four hyperthreads, we run the following strategies.
%  N ==5e5
\begin{itemize}
    \item Sequentiel: and here we just compute the sum of the first $N$ numbers without any concurency
    \item Multithreading: and here we use the \texttt{ ThreadPoolExecutor} library to compute the sum of the first $N$ numbers, We devide $N$ by the number of threads and each thread will compute the sum of the numbers in its range. For our case we used 4 threads.
    \item multiprocessing : and in this case, we use the \texttt{ProcessPoolExecutor} library to calculate the sum of the first $N$ numbers. We divide $N$ by the number of processes, and each process sums the numbers in its range. In this situation, we employed 2,4,8,16 processes, correspondingly.
\end{itemize}

Figure~\ref{fig:python_multiprocessing_strategies} shows the results of the experiment. As one can notice
\begin{figure}[!hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_multiprocessing}
    \caption{Energy behaviour resulting from diffrent python concurrency strategies (number of workers) }
    \label{fig:python_multiprocessing_strategies}
\end{figure}

Additionally, \Cref{fig:python_multiprocessing} reports on the energy consumption of Python applications when introducing parallelism.
Python applications are single-threaded systems constrained by the \emph{global lock system} (GLS).
However, due to the increase of cores/threads per CPU, many Python libraries started to take advantage of this hardware feature by allowing concurrent execution of Python processes.
For example, the multiprocessing library\footnote{\url{https://docs.python.org/3/library/multiprocessing.html}} spawns sub-processes to increase the degree of concurrency of Python applications.
As one can see in \Cref{fig:python_multiprocessing}, the energy consumption is correlated with the number of threads until reaching the limit of physical cores, when concurrent processes start to compete for the CPU.

@@

\clearpage
\input{python_interperters.tex}
% \subsection{threads to validity \note{missing}}
\clearpage
\section{Summary}
Because Python is widely used by software services deployed in public and private cloud infrastructures, increasing the energy efficiency of Python-based applications will reduce ICT carbon emissions.
This chapter discusses various ways for increasing the energy efficiency of Python-based applications. It first explains why such a choice was made, and then it develops a method to enhance the energy efficiency of the most commonly employed use cases, which are revealed to include web development and machine learning.
In web development, we found that the majority of the energy is wasted while processing requests, therefore adopting the proper strategy can reduce energy consumption by tenfold.
Following that, we looked at the energy usage for the machine learning usecase and discovered that it is not proportional to model accuracy.
As a result, a reduction in accuracy can result in large energy savings.
Following that, we dug deeper to investigate how data structures, parallelism, and iterative methods affect energy use.
In the second section, we presented a simple technique for optimizing the energy usage of Python-based apps. without making substantial changes to the code This is the use of many Python interpreters
Section 2 started by categorizing and filtering these implementations into three major kinds (compiler, interpreter, and extra libraries).
Then we conducted a series of experiments to compare the energy consumption of various interpreters.
The findings indicate the lack of a general solution and the importance of selecting the appropriate interpreter for the appropriate use case. In certain circumstances, some of these answers may be the best, while in others, they may be the worst.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.
We believe that our contributions will benefit a wide spectrum of legacy systems, not only reducing ICT's carbon footprint, but also lowering their cloud bills for the resources these services need.
